Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-26 article.tex
section 4 suite

In the bottom part of Table \ref{tab:comparison}, the proposed
approach is compared with three other methods: standard (Pearson)
correlation (PC), generalized transfer entropy (GTE), and GENIE3. ROC
and PR curves on the \textit{normal-2} network are shown for all
methods in Figure \ref{fig:}. Pearson correlation measures the
unconditional linear (in)dependence between variables and it should
thus not be able to filter out indirect interactions between
neurons. GTE \citep{stetter2012model} was proposed as a baseline for
the challenge. This method builds on Transfer Entropy to measure the
association between two neurons. Unlike our approach, it can predict
the direction of the edges. GENIE3 \citep{huynhthu2010inferring} is a
gene regulatory network inference method that was the best performer
of the DREAM5 challenge \cite{dream5paper}. When transposed to neural
networks, this method uses as a confidence score for the edge going
from neuron $i$ to neuron $j$ the importance score of variable $X_i$
in a Random Forests model trying to predict $X_j$ from all variables
in $X\setminus X_j$. To reduce computing times of this method, we had
to reduce each tree in the Random Forests model to a maximum depth of
3. This constraint potentially severly affects the performance of this
method with respect to the use of fully grown trees. We nevertheless
provide these results for comparison purpose. PC and GENIE3 were both
applied on the same pre-filtered time series as partial
correlation. For GENIE3, we built 10000 trees per neuron and we used
default settings for all other parameters (except for the maximal tree
depth). For GTE, we reproduced the exact same setting (conditioning
level and pre-processing) that was used for the challenge.

@@

Partial correlation clearly outperforms all other methods. The improvement is more important in terms of AUPRC bu by about a factor
of 2 in AUPRC reaching $0.347$ and an increased of $0.03$ in
AUROC. Averaging Partial correlation matrix over a wide of parameters
further improves AUPRC to $0.389$ and improve AUROC by $0.009$.  Two
reasons might explain such differences: (i) only the partial
correlation-based methods and GENI3 at lesser extent distinguish
direct links from indirect one, (ii) simulated networks from the
\citep{stetter2012model} generator or neural networks in general could
be reduced to a linear model after pre-processing. Those results
generalizes over the other datasets see Appendix \textcolor{red}{REF}.

>>>>

In the bottom of Table \ref{tab:comparison}, we provide as a comparison the
performance of three other methods: standard (Pearson) correlation (PC),
generalized transfer entropy (GTE), and GENIE3. ROC and PR curves on the
\textit{normal-2} network are shown for all methods in Figure
\ref{fig:}. Pearson correlation measures the unconditional linear
(in)dependence between variables and it should thus not be able to filter out
indirect interactions between neurons. GTE \citep{stetter2012model} was
proposed as a baseline for the challenge. This method builds on Transfer
Entropy to measure the association between two neurons. Unlike our approach, it
can predict the direction of the edges. GENIE3 \citep{huynhthu2010inferring} is
a gene regulatory network inference method that was the best performer of the
DREAM5 challenge \cite{dream5paper}. When transposed to neural networks, this
method uses as a confidence score for the edge going from neuron $i$ to neuron
$j$ the importance score of variable $X_i$ in a Random Forests model trying to
predict $X_j$ from all variables in $X\setminus X_j$. To reduce computing times
of this method, we had to reduce each tree in the Random Forests model to a
maximum depth of 3. This constraint potentially severly affects the performance
of this method with respect to the use of fully grown trees. We nevertheless
provide these results for comparison purpose. PC and GENIE3 were both applied
on the same pre-filtered time series as partial correlation. For GENIE3, we
built 10000 trees per neuron and we used default settings for all other
parameters (except for the maximal tree depth). For GTE, we reproduced the
exact same setting (conditioning level and pre-processing) that was used for
the challenge.

@@

Partial correlation and average partial correlation clearly outperforms all
other methods. The improvement is more important in terms of AUPRC than in
terms of AUROC. As expected, Pearson correlation performs very poorly in terms
of AUPRC. GTE and GENIE3 works much better but these two methods are
nevertheless clearly below partial correlation. Among these two methods, GTE is
slightly better in terms of AUROC, while GENIE3 is significantly better in
terms of AUPRC. Given that we had to limit this latter method for computational
reasons, these results are very promising and a comparison with the full GENIE3
approach is certainly part of our future works.

%% bu by about a factor
%% of 2 in AUPRC reaching $0.347$ and an increased of $0.03$ in
%% AUROC. Averaging Partial correlation matrix over a wide of parameters
%% further improves AUPRC to $0.389$ and improve AUROC by $0.009$.  Two
%% reasons might explain such differences: (i) only the partial
%% correlation-based methods and GENI3 at lesser extent distinguish
%% direct links from indirect one, (ii) simulated networks from the
%% \citep{stetter2012model} generator or neural networks in general could
%% be reduced to a linear model after pre-processing. Those results
%% generalizes over the other datasets see Appendix \textcolor{red}{REF}.
